\chapter*{About MATLAB}

\section{What is \mlab?}
\mlab is produced by MathWorks, and is one of a number of commercially available software packages for numerical computing and programming. \mlab\ provides an interactive environment for algorithm development, data visualisation, data analysis, and numerical computation. \mlab, which derives its name from MATrix LABoratory, excels at matrix operations and graphics. Its main competitors are Maple, Mathematica, and Mathcad, each with their own strengths and weaknesses. 

\graffito{\mlab student version is available for around \pounds55 + VAT}\mlab is available in both commercial and academic versions with new releases binannually \eg\ R2016a (released around March 2016), and R2016b (released around September 2016). \mlab itself is the core product and is augmented by additional toolboxes, many of which have to be purchased separately. If you want to run \mlab on your own computer MathWorks offers a student version (\href{http://www.mathworks.com/academia/student_version/}{http://www.mathworks.com/academia/student\_version/}) with some of the most commonly used toolboxes for around \pounds55 +VAT. The accompanying online material, and the first version of screenshots for this document were originally based on the R2009a release of \mlab. As the \mlab user interface was substantially redeveloped by Mathworks few years later the current version of the course has been redesigned to conform to this change and is now based on the new User interface of \mlab R2013a running under Microsoft Windows. \mlab is available for other operating systems such are Mac OS X and Linux. The User Interface is very similar and the material in this course is directly applicable to versions of \mlab running under these operating systems. Finally, as the course aims to present very basic concepts of \mlab the material is widely applicable to most versions of \mlab.

\section{How is \mlab used in industry?}
\graffito{Knowing how to use \mlab is a vital skill for many engineering jobs!}The ability to use tools such as \mlab is increasingly required by employers of graduate engineers in industry. Many job adverts specifically mention knowledge of \mlab as an essential skill.

\mlab is a widely-used tool in many different fields of engineering and science. The following is a brief list of examples from Chemical, Civil, Electrical, and Mechanical Engineering:
\begin{itemize}
\item Motorsport Teams Improve Vehicle Performance with MathWorks Tools\\
(\href{http://uk.mathworks.com/company/user_stories/penske-technology-group-enables-motorsport-teams-to-improve-vehicle-performance.html}{http://uk.mathworks.com/company/user\_stories/penske-technology-group-enables-motorsport-teams-to-improve-vehicle-performance.html})
\item Bell Helicopter Develops the First Civilian Tiltrotor\\
(\href{http://www.mathworks.com/company/newsletters/news_notes/oct06/bellhelicopter.html}{http://www.mathworks.com/company/newsletters/news\_notes/oct06/bellhelicopter.html})
\item Modeling Flexible Bodies in SimMechanics and Simulink\\
(\href{http://uk.mathworks.com/company/newsletters/articles/modeling-flexible-bodies-in-simmechanics-and-simulink.html?s_tid=srchtitle}{http://uk.mathworks.com/company/newsletters/articles/modeling-flexible-bodies-in-simmechanics-and-simulink.html?s\_tid=srchtitle})
%\item Greenhouse Designed with \mlab and Simulink Revolutionizes Agriculture in Arid Coastal Regions\\
%(\href{http://www.mathworks.com/company/user_stories/userstory2347.html?by=industry}{http://www.mathworks.com/company/user\_stories/userstory2347.html?by=industry})
%\item Thames Water Aims to Reduce Leaks by More Than 25\% Using a MATLAB-Based Leak-Location System\\
%(\href{http://www.mathworks.com/company/user_stories/userstory2354.html?by=industry}{http://www.mathworks.com/company/user\_stories/userstory2354.html?by=industry})
\item Samsung UK Develops 4G Wireless Systems with Simulink\\
(\href{http://uk.mathworks.com/company/user_stories/samsung-uk-develops-4g-wireless-systems-with-simulink.html?s_tid=srchtitle}{http://uk.mathworks.com/company/user\_stories/samsung-uk-develops-4g-wireless-systems-with-simulink.html?s\_tid=srchtitle})
\item Elektrobit Testing Ltd. Develops High-Resolution Radio Channel Measuring System\\
(\href{http://uk.mathworks.com/company/user_stories/elektrobit-testing-ltd-develops-high-resolution-radio-channel-measuring-system.html?s_tid=srchtitle}{http://uk.mathworks.com/company/user\_stories/elektrobit-testing-ltd-develops-high-resolution-radio-channel-measuring-system.html?s\_tid=srchtitle})
\item Halliburton Makes Oil Exploration Safer Using MATLAB and Neural Networks\\
(\href{http://uk.mathworks.com/company/user_stories/halliburton-makes-oil-exploration-safer-using-matlab-and-neural-networks.html?s_tid=srchtitle}{http://uk.mathworks.com/company/user\_stories/halliburton-makes-oil-exploration-safer-using-matlab-and-neural-networks.html?s\_tid=srchtitle})
\end{itemize}